Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-05-05 SIR_manuscript.tex
analysis, R0 wrt rho

\begin{multline}
\label{eq:a3}
a_3 = \frac{w_I}{w_S}
\Bigg(\Big(
( \theta_{{w}}{{\it p\_I}}^{2}-\theta_{{c}}{{\it p\_I}}^
{2}\theta_{{w}} ) {\omega}^{3} 
+ ( 2\,\theta_{{w}}{{\it p\_I
}}^{2}-\theta_{{c}}{{\it p\_I}}^{2}-{\it p\_I}\,{\theta_{{w}}}^{2}-
\theta_{{c}}{\it p\_I}\,\theta_{{w}}-{{\it p\_I}}^{2}{\theta_{{w}}}^{2
}+2\,\theta_{{w}}{\it p\_I} ) \gamma\,{\omega}^{2} \\
 + ( -2\,{ \it p\_I}\,{\theta_{{w}}}^{2}-\theta_{{c}}{\it p\_I}+3\,\theta_{{w}}{
\it p\_I}-{\theta_{{w}}}^{2}+\theta_{{w}} ) {\gamma}^{2}\omega 
 + ( \theta_{{w}}-{\theta_{{w}}}^{2} ) {\gamma}^{3} \Big) \frac{w_I}{w_S}
 + ( -\theta_{{c}}{\it p\_I}\,\theta_{{w}}+\theta_{{c}}{\it p\_I}
 ) \gamma\,{\omega}^{2}+ \\
 ( \theta_{{w}}+\theta_{{c}}{\it
p\_I}-{\theta_{{w}}}^{2}-\theta_{{c}}{\it p\_I}\,\theta_{{w}} )
{\gamma}^{2}\omega+ ( \theta_{{w}}-{\theta_{{w}}}^{2} ) {
\gamma}^{3}
\Bigg),
\end{multline}
\begin{align}
\label{eq:abcd2}
b_3&= 2\frac{w_I}{w_S}(\omega+\gamma)\gamma\Big( 
(\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w 
\Big) ,\\
c_3&= (\omega+\gamma)^2\,\gamma\Big( 
(2-\theta\_w)\,\gamma^2\,\theta\_w+
(1+\frac{w_I}{w_S})\,\omega\,\gamma\,\theta\_w+
\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
d_3&= \frac{(\omega+\gamma)}{\omega} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
\end{align}
Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$. 
If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$. An example of this scenario occures when $\theta\_w=0$ and $\theta\_c=1$, which is illustraited in the top-right panel of the Fig.\,\ref{pan2} panel (b). This is when the strength of isolation for awaiting people is the least, but the most for the confirmed cases. In this case 
\ali{I stoped here!}
\begin{equation}
\label{ineq:a3}
a_3 <0\iff\omega+\gamma<(\omega\,p_I+\gamma)\frac{w_I}{w_S}.
\end{equation}
Given that $\frac{w_I}{w_S}\geq 1$, and $0\leq\,p_I\leq\,1$, this inequality is hold when a perfect sensitive test is used, i.e., $p_I=1$, and a targeted tesing strategy is followed, i.e., $\frac{w_I}{w_S}>1$. \ali{i)what is the interpretation of this? 2) (Ben:) calculate the corresponding eigenvector to $\Rnum$, scale it to 1, and plot the population of each infected compartment, what does this tell us? }

Note that $\rho$ and $\omega$ have similar mechanism in delaying people to get into $I\_c$. The non monotonic countereffect of testing rate and test return rate are not obvious. 
that people in the DFE waiting for negative test, $S_n^*$.
More testing may have a counter effect on $\Rnum$, as if we have more testing we have more people in $S_n^*$ which will take them longer to get to $I\_c$.

more $\rho$ then

>>>>

\begin{multline}
\label{eq:a3}
a_3 = \frac{w_I}{w_S}
\Bigg(\Big(
( \theta_{{w}}{{\it p\_I}}^{2}-\theta_{{c}}{{\it p\_I}}^
{2}\theta_{{w}} ) {\omega}^{3} 
+ ( 2\,\theta_{{w}}{{\it p\_I
}}^{2}-\theta_{{c}}{{\it p\_I}}^{2}-{\it p\_I}\,{\theta_{{w}}}^{2}-
\theta_{{c}}{\it p\_I}\,\theta_{{w}}-{{\it p\_I}}^{2}{\theta_{{w}}}^{2
}+2\,\theta_{{w}}{\it p\_I} ) \gamma\,{\omega}^{2} \\
 + ( -2\,{ \it p\_I}\,{\theta_{{w}}}^{2}-\theta_{{c}}{\it p\_I}+3\,\theta_{{w}}{
\it p\_I}-{\theta_{{w}}}^{2}+\theta_{{w}} ) {\gamma}^{2}\omega 
 + ( \theta_{{w}}-{\theta_{{w}}}^{2} ) {\gamma}^{3} \Big) \frac{w_I}{w_S}
 + ( -\theta_{{c}}{\it p\_I}\,\theta_{{w}}+\theta_{{c}}{\it p\_I}
 ) \gamma\,{\omega}^{2}+ \\
 ( \theta_{{w}}+\theta_{{c}}{\it
p\_I}-{\theta_{{w}}}^{2}-\theta_{{c}}{\it p\_I}\,\theta_{{w}} )
{\gamma}^{2}\omega+ ( \theta_{{w}}-{\theta_{{w}}}^{2} ) {
\gamma}^{3}
\Bigg),
\end{multline}
\begin{align}
\label{eq:abcd2}
b_3&= 2\frac{w_I}{w_S}(\omega+\gamma)\gamma\Big( 
(\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w 
\Big) ,\\
c_3&= (\omega+\gamma)^2\,\gamma\Big( 
(2-\theta\_w)\,\gamma^2\,\theta\_w+
(1+\frac{w_I}{w_S})\,\omega\,\gamma\,\theta\_w+
\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
d_3&= \frac{(\omega+\gamma)}{\omega} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
\end{align}
Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$. 
If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$. An example of this counter-vailing effect of $\rho$ on $\Rnum$ occures when $\theta\_w=0$ and $\theta\_c=1$. This is illustraited in the top-right panel of the Fig.\,\ref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases. In this case $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
Specifically, in the case of targeted testing which is identified with $\frac{w_I}{w_S}> 1$, and using a perfect sensative test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.  
Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the counter-vailing effect, which is not obvious, of these two parameters on $\Rnum$. 
% that people in the DFE waiting for negative test, $S_n^*$.
% More testing may have a counter effect on $\Rnum$, as if we have more testing we have more people in $S_n^*$ which will take them longer to get to $I\_c$.
\ali{I stoped here!}
